---
title: "SI 618 Homework 3"
author: "Nishan Bose"
date: "Wednesday 01 April 2015"
output: html_document
---

```{r, echo=FALSE}
# Connecting to an sqlite Database
# Load the libraries needed for this
library(DBI)
library(RSQLite)
library(plyr)
library(ggplot2)

# Load the driver needed for this
driver<-dbDriver("SQLite")

# Connect to the database "cars"
connect<-dbConnect(driver, dbname = "vehicles.db")

# Loading information into data frames
carsTable = dbGetQuery(connect, "select * from vehicles")
```

##Getting Data from SQLite Database

In the data preparation step, a SQLite database has been created and populated with vehicle data. Now the data in the database is retrieved using R package DBI and RSQLite and stored in a data frame named vehicles. Here are the first 10 rows of the data frame, and the summary.

```{r, fig.width = 16, fig.height=5, dpi=128, echo=FALSE}
#Displaying first ten entries
head(carsTable,n=10)

#Displaying summary
summary(carsTable)
```

##Converting to Factors

To make downstream analysis easier, we convert the data in columns vehicles$make, vehicles$vclass, vehicles$cylinders, and vehicles$trany into factors. Here is the summary of the data frame after the conversion.

```{r, fig.width = 16, fig.height=5, dpi=128, echo=FALSE}
#Converting into factors
carsTable$make <- as.factor(carsTable$make)
carsTable$vclass <- as.factor(carsTable$vclass)
carsTable$cylinders <- as.factor(carsTable$cylinders)
carsTable$trany <- as.factor(carsTable$trany)
summary(carsTable)
```

##Filter Down Data

We will filter down the data such that only ‘VClass’ with more than 40 vehicles are kept. Here is the summary of the data frame after this subsetting step.

```{r, fig.width = 16, fig.height=5, dpi=128, echo=FALSE}
#Filtering down data
carsTable.ddply <- ddply(carsTable, c("vclass"), transform, vclass.size = length(vclass))
carsTable.subset <- subset(carsTable.ddply, vclass.size > 40)
carsTable.subset <- subset(carsTable.subset, select = -vclass.size)
summary(carsTable.subset)
```

##Fuel Economy of Vehicles of Different Makes

For each vehicle class in filtered down data, we plot the mean combined MPG (average of data in vehicles$comb08) for each vehicle maker every year. And then, we compute at the mean combined MPG in all years for each vehicle maker, and plot it. Both charts are created with ggplot(). Note how the vehicle makers are ranked in the second plot.

```{r, fig.width = 16, fig.height=5, dpi=128, echo=FALSE}
vclass.unique <- unique(carsTable.subset$vclass)

for (item in vclass.unique) {
  vclass.subset <- subset(carsTable.subset, vclass==item)
  
  year.data <- ddply(vclass.subset, c("make","year"), summarise, avg.data = mean(combo08))
  year.plot <- ggplot(year.data, aes(year, avg.data)) + ggtitle(item) + xlab("Year") + ylab("Mean Combined MPG") + guides(col=guide_legend(ncol=2)) + geom_line(aes(colour = make))
  print(year.plot)
  
  make.data <- ddply(vclass.subset, c("make"), summarise, avg.data = mean(combo08))
  make.plot <- qplot(reorder(make, avg.data, function(x) -x), y = avg.data, data = make.data, geom="bar", stat = "identity", alpha = I(0.5), main=(item)) + xlab("Make") + ylab("Mean Combined MPG in All Years") + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=1));
  print(make.plot)
}
```